---
title: "Coursera's Machine Learning Course Assignment"
output: html_document
---
   
   
### Background
    
Data were collected during a weight lifting exercise using accelerometer data on the belt, forearm, arm, and dumbbell of 6 participants. They were told to do the exercise in five different ways, only one of which was correct.
The aim of this assignment is to find a machine learning algorithm that should help to identify the five different ways (labeled 'A' to 'E') by using the data from http://groupware.les.inf.puc-rio.br/har. 
   
### Data
   
```{r}   
##First of all, the working directory is set

```
   
The necessary libraries ("caret", "randomForest") are loaded:
```{r}
library(caret)
library(randomForest)
```
Data are now collected from Coursera's website.
  
First, the original training-data are downloaded into the working directory:  
```{r}  
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
```
  
Second, the original test-data are downloaded as well:   
```{r}
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
```
  
### Cleaning Data  
Both files are read into data frames, simultaneously converting empty values to NA's.
  
```{r}  
trainData <- read.csv("pml-training.csv", header = T, na.strings= c("", NA))
testData <- read.csv("pml-testing.csv", header = T, na.strings= c("", NA))
```
   
The original training-data set consists of 160 columns.
   
```{r}
ncol(trainData)
```
The first 7 columns contain meaningless information for the purpose of this study; only the remaining columns are transferred to a reduced data set.
  
```{r}
reducedData <- trainData[, 8:160]

```
Furthermore, there are many columns with lots of NA's - all columns with NA's are deleted. The remaining 53 columns are the subject for the following procedures.
   
```{r}
reducedData <- reducedData[,  colSums(is.na(reducedData))==0]
ncol(reducedData)
```
  
### Data Partition
  
The reduced data set is partitioned into a training- (60%) and test-set.
  
```{r}
train <- createDataPartition(y=reducedData$classe, p=0.6, list=F)
training <- reducedData[train, ]
testing <- reducedData[-train, ]
```
  

### The model: Random Forest
  
Due to the character of the features and the wish for high accuracy, Random Forest was taken as the method of choice for this study.  
However, all attempts to use the train-function for this aim failed.
Instead, the genuine randomForest-function succeeded immediately. To guarantee reproducibility, the seed was set first.
  
```{r}
set.seed(1234)
modelFit = randomForest(classe ~., data=training)
modelFit
```
  
```{r}  
pred <- predict(modelFit, testing)
confusionMatrix(pred, testing$classe)
```
   
With an **out-of-sample accuracy** of over **99.3%**, the model proves to be highly accurate.
This is underlined by the results of the model on the 20 test-cases, all of which were classified accurately.
  
```{r}  
results = predict(modelFit, testData)
results
```

